---
title: "ChatGPT Comparison Analysis"
author: "Monnie McGee"
date: "2024-04-24"
output: html_document
editor_options: 
  chunk_output_type: console
---

The following assumes that the data file `AllQuestionsFinal2.csv` is locating in the current working directory for R. Use `getwd()` to determine the current working directory. If a change is needed, use `setwd()`.

Ordinal logistic regression was used to analyze the results. 

```{r Packages needed}
library(tidyverse) # data manipulation
library(ggplot2) # graphics
library(VGAM) # ordinal logistic regression
library(MASS) # ordinal logistic regression (double check)
library(brant) # check proportionl odds ratio
```

```{r Enter data}

## dfWide = data in wide format (2 columns for scores)
## dfLong = data in long format (1 column for scores)
dfWide <- read.csv(file.choose(), header=T)
dfWide$Score4_0 = if_else(dfWide$Score4_0 == 5,4,as.numeric(dfWide$Score4_0))
dfWide$Score3_5 = if_else(dfWide$Score3_5 == 5,4,as.numeric(dfWide$Score3_5))
dfWide$Image <- as.factor(dfWide$Image)
dfWide$Type <- as.factor(dfWide$Type)
dfWide$Exam <- as.factor(dfWide$Exam)
dfWide$Diff <- dfWide$Score4_0 - dfWide$Score3_5
table(dfWide$Diff) 
```

The table shows 8 different levels, some with only 1 observation. Therefore, we rescale the `Diff` variable into a variable called `Compare`. The rescaling is given by the following heuristic:

- if Diff < 0, ChatGPT4.0 is worse
- if Diff = 0, ChatGPT3.5 and 4.0 are the same
- if Diff = 1 or 2, GPT4.0 is marginally better
- if Diff > 2 GPT4.0 is much better

Free response and multiple choice questions were coded together.  For MC questions, when GPT 4 wins, the score is 3. This is fine. On a test, MC questions are generally all correct or all incorrect. Therefore, missing a MC question is equivalent to getting partial  credit on a FR question. 

```{r Rescaling}
dfWide <- dfWide %>% mutate(
  Compare = case_when(
    Diff < 0 ~ 'Worse',
    Diff == 0 ~ 'Same',
    Diff == 1 | Diff == 2 ~ "Somewhat",
    Diff > 2 ~ "Much"
    )
 )
#dfWide$Compare <- ordered(dfWide$Compare, levels = c("Worse","Same","Somewhat","Much") ) # make Compare an ordered factor and specify order
dfWide$Compare <- ordered(dfWide$Compare, levels = c("Much","Somewhat", "Same", "Worse") ) # make Compare an ordered factor and specify order

str(dfWide)
table(dfWide$Compare)
```

Numbers within categories are still somewhat lopsided, but they are at least better. No category has only one entry.

Creating the difference variable is a way of taking the paired nature of the questions into account.

Use MASS library

```{r OrdinalLogReg}
## Compare is the response. Takes platform into account.
## Type = FR or MC is an explanatory variable
## Image = 0 or 1 means an image was involved (0 = No, 1 = yes)
fit.polr <- polr(Compare ~ Image + Type, data=dfWide)
summary(fit.polr)
brant(fit.polr)
```

Now use VGAM library for comparison.
```{r OridinalLogRegVGAM}
dfWide$Type = relevel(dfWide$Type, ref = "MC")
fit.vglm <- vglm(Compare ~ Type+Image, data=dfWide, cumulative(parallel=TRUE))
#fit.vglm <- vglm(Compare ~ Exam, data=dfWide, cumulative(parallel=TRUE))
summary(fit.vglm)
QType = data.frame(Type = "MC", Image = "0")
predict(fit.vglm, type = "response", newdata = QType)
QType = data.frame(Type = "MC", Image = "1")
predict(fit.vglm, type = "response", newdata = QType)
QType = data.frame(Type = "FR", Image = "0")
predict(fit.vglm, type = "response", newdata = QType)
QType = data.frame(Type = "FR", Image = "1")
predict(fit.vglm, type = "response", newdata = QType)

#exp(confint(fit.vglm))


# Estimate individual probabilities

#PCMuch)
PMuch_MC = exp(-1.8256) / (1 + exp(-1.8256))
PMuch_MCI = exp(-1.8256 +1.9172) / (1+exp(-1.8256 +1.9172))
PMuch_FR = exp(-1.8256 +.4565) / (1+ exp(-1.8256 +.4565))
PMuch_FRI = exp(-1.8256 +.4565+1.9172) / (1+exp(-1.8256 +.4565+1.9172))


#P(Somewhat) = P(X <= Somewhat) - P(X = Much)
PSomewhat_MC = (exp(-1.1350) / (1 + exp(-1.1350))) - (exp(-1.8256) / (1 + exp(-1.8256)))
PSomewhat_MCI = (exp(-1.1350 +1.9172) / (1+exp(-1.1350 +1.9172))) - (exp(-1.8256 +1.9172) / (1+exp(-1.8256 +1.9172)))
PSomewhat_FR = (exp(-1.1350 +.4565) / (1+ exp(-1.1350 +.4565))) - (exp(-1.8256 +.4565) / (1+ exp(-1.8256 +.4565)))
PSomewhat_FRI = (exp(-1.1350 +.4565+1.9172) / (1+exp(-1.1350 +.4565+1.9172))) - (exp(-1.8256 +.4565+1.9172) / (1+exp(-1.8256 +.4565+1.9172)))

#P(Same) = P(X <= Same) - P(X <= Somewhat)
PSame_MC = (exp(2.2053 ) / (1 + exp(2.2053 ))) - ((exp(-1.1350) / (1 + exp(-1.1350))))
PSame_MCI = (exp(2.2053  +1.9172) / (1+exp(2.2053  +1.9172))) - (exp(-1.1350 +1.9172) / (1+exp(-1.1350 +1.9172)))
PSame_FR = (exp(2.2053  +.4565) / (1+ exp(2.2053  +.4565))) - ((exp(-1.1350 +.4565) / (1+ exp(-1.1350 +.4565))))
PSame_FRI = (exp(2.2053  +.4565+1.9172) / (1+exp(2.2053  +.4565+1.9172))) - (exp(-1.1350 +.4565+1.9172) / (1+exp(-1.1350 +.4565+1.9172)))

#P(Worse) = 1 - P(X <= Same)
PWorse_MC = 1 - (exp(2.2053 ) / (1 + exp(2.2053 )))
PWorse_MCI = 1 - (exp(2.2053  +1.9172) / (1+exp(2.2053  +1.9172)))
PWorse_FR = 1 - (exp(2.2053  +.4565) / (1+ exp(2.2053  +.4565)))
PWorse_FRI = 1 - (exp(2.2053  +.4565+1.9172) / (1+exp(2.2053  +.4565+1.9172)))




```

# Plot Creation

To create a spaghetti plot, we need the data in long format.
```{r PivotLonger}
dfLong <- dfWide %>% pivot_longer(Score3_5:Score4_0,
                      names_to=c("Mark","Platform"),
                      values_to="Score", names_sep=5)
set.seed(1)
dfLong$ScoreJitter = dfLong$Score + rnorm(dim(dfLong)[1],0,.1) # Jitter
dfLong$Exam <- as.factor(dfLong$Exam)
dfLong$Type <- as.factor(dfLong$Type)
dfLong$Image <- as.factor(dfLong$Image)
dfLong$Platform <- as.factor(dfLong$Platform)
str(dfLong)
```

Create a spaghetti plot where line type indicates presence of image. MC and FR questions are used.
```{r Plots}
# Create the ggplot
p1 <- ggplot(data = dfLong, aes(x = Platform, y = ScoreJitter, group = Question)) +
    mytheme +
    scale_colour_grey(start = 0.4, end = 0.6) +  # Set the grey color scale for the lines
    labs(
        y = "Score",  # y-axis label
        x = "Platform"  # x-axis label
    ) +
    geom_line(lwd = 0.5) +  # Set the line width
    scale_x_discrete(labels = c("ChatGPT3.5", "ChatGPT4")) +  # Custom labels for x-axis
    ylim(-0.25, 4.5)  # Set limits for the y-axis

# Print the plot
print(p1)
```


Create a spaghetti plot where line type indicates presence of image. MC and FR questions are used. Remove the concordant lines.


```{r Plots}

# Uncomment to reemove Concordant lines
dfLong <- dfLong %>% filter(Diff != 0)

# Create the ggplot
p1 <- ggplot(data = dfLong, aes(x = Platform, y = ScoreJitter, group = Question)) +
    mytheme +
    scale_colour_grey(start = 0.4, end = 0.6) +  # Set the grey color scale for the lines
    labs(
        y = "Score",  # y-axis label
        x = "Platform"  # x-axis label
    ) +
    geom_line(lwd = 0.5) +  # Set the line width
    scale_x_discrete(labels = c("ChatGPT3.5", "ChatGPT4")) +  # Custom labels for x-axis
    ylim(-0.25, 4.5)  # Set limits for the y-axis

# Print the plot
print(p1)
```


## Spaghetti Plot for McNemar

Here we code an Incorrect answer as 0 and a correct answer as one
```{r PivotLonger}

dfWide$Correct3_5 = if_else(dfWide$Score3_5 >= 3,3,0)
dfWide$Correct4_0 = if_else(dfWide$Score4_0 >= 3,3,0)
dfWide$ConOrDis = if_else(dfWide$Correct3_5 == dfWide$Correct4_0,"Concordant","Discordant")

#For plottling all pairs
dfLong <- dfWide %>% pivot_longer(Correct3_5:Correct4_0,
                      names_to=c("Mark","Platform"),
                      values_to="Score", names_sep=7)

set.seed(1)
dfLong$ScoreJitter = dfLong$Score + rnorm(dim(dfLong)[1],0,.1) # Jitter

dfLong$Exam <- as.factor(dfLong$Exam)
dfLong$Type <- as.factor(dfLong$Type)
dfLong$Image <- as.factor(dfLong$Image)
dfLong$Platform <- as.factor(dfLong$Platform)
str(dfLong)
```


```{r Plots}

# Uncomment to reemove Concordant lines
dfLong <- dfLong %>% filter(ConOrDis == "Discordant")

# Create the ggplot
p1 <- ggplot(data = dfLong, aes(x = Platform, y = ScoreJitter, group = Question)) +
    mytheme +
    scale_colour_grey(start = 0.4, end = 0.6) +  # Set the grey color scale for the lines
    labs(
        y = "Score",  # y-axis label
        x = "Platform"  # x-axis label
    ) +
    geom_line(lwd = 0.5) +  # Set the line width
    scale_x_discrete(labels = c("ChatGPT3.5", "ChatGPT4")) +  # Custom labels for x-axis
    ylim(-0.25, 3.5)  # Set limits for the y-axis

# Print the plot
print(p1)
```
