---
title: "Equity Project"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## references used in write up: https://web.pdx.edu/~newsomj/ho_chisq%20within.pdf
## Read in the Data


```{r}
library(tidyverse)
df = read.csv("/Users/26438728/Library/CloudStorage/Box-Box/Monnie and Bivin Paper 2024/Raw Tests and Results/Results/Final Results/AllQuestionsFinalMcGee.csv", stringsAsFactors = TRUE)
#df = read.csv(file.choose(),header = TRUE)
df = df %>% filter(Exam != "SMUExam")
```

## Create Correct / Incorrect Columns

```{r}
df$Correct3_5 = if_else(df$Score3_5 >= 3,1,0)
df$Correct4_0 = if_else(df$Score4_0 >= 3,1,0)
```

## Contingency Table to Mcnemar's Test Overall
```{r}

PP = sum(if_else((df$Correct3_5 == "1" & df$Correct4_0 == "1"),1,0))
NP = sum(if_else((df$Correct3_5 == "0" & df$Correct4_0 == "1"),1,0))
PN = sum(if_else((df$Correct3_5 == "1" & df$Correct4_0 == "0"),1,0))
NN = sum(if_else((df$Correct3_5 == "0" & df$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT

```

## McNemar's Test Overall
```{r}

# 3.5 versus 4.0 for all the data
mcnemar.test(CT)

```




## Contingency Table to McNemar's Test by Image
```{r}

# Filter for Each Question Type 
df_Image = df %>% filter(Image == "1") # Has an Image
df_No_Image = df %>% filter(Image == "0") # Does not have an Image
  
#Image
PP = sum(if_else((df_Image$Correct3_5 == "1" & df_Image$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_Image$Correct3_5 == "0" & df_Image$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_Image$Correct3_5 == "1" & df_Image$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_Image$Correct3_5 == "0" & df_Image$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_Image = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_Image

#No Image
PP = sum(if_else((df_No_Image$Correct3_5 == "1" & df_No_Image$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_No_Image$Correct3_5 == "0" & df_No_Image$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_No_Image$Correct3_5 == "1" & df_No_Image$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_No_Image$Correct3_5 == "0" & df_No_Image$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_No_Image = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_No_Image

```

## McNemar's Test by Image
```{r}

# 3.5 versus 4.0

#Image
mcnemar.test(CT_Image)
#No Image
mcnemar.test(CT_No_Image)

```






## Contingency Table to Mcnemar's Test by Question Type
```{r}

# Filter for Each Question Type 
df_FR = df %>% filter(Type == "FR") 
df_MC = df %>% filter(Type == "MC") 
  


#FR
PP = sum(if_else((df_FR$Correct3_5 == "1" & df_FR$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_FR$Correct3_5 == "0" & df_FR$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_FR$Correct3_5 == "1" & df_FR$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_FR$Correct3_5 == "0" & df_FR$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_FR = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_FR


#MC
PP = sum(if_else((df_MC$Correct3_5 == "1" & df_MC$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_MC$Correct3_5 == "0" & df_MC$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_MC$Correct3_5 == "1" & df_MC$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_MC$Correct3_5 == "0" & df_MC$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_MC = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_MC

```

## McNemar's Test by Question Type
```{r}

# 3.5 versus 4.0

#FR
mcnemar.test(CT_FR)
#MC
mcnemar.test(CT_MC)

```



## Contingency Table to Mcnemar's Test by Question Type and Image
```{r}

# Filter for Each Question Type
df_FR_NoI = df %>% filter(Type == "FR" & Image == "0") 
df_FR_I = df %>% filter(Type == "FR" & Image == "1") 
df_MC_NoI = df %>% filter(Type == "MC" & Image == "0") 
df_MC_I = df %>% filter(Type == "MC" & Image == "1") 
  


#FR_NoI
PP = sum(if_else((df_FR_NoI$Correct3_5 == "1" & df_FR_NoI$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_FR_NoI$Correct3_5 == "0" & df_FR_NoI$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_FR_NoI$Correct3_5 == "1" & df_FR_NoI$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_FR_NoI$Correct3_5 == "0" & df_FR_NoI$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_FR_NoI = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_FR_NoI


#FR_I
PP = sum(if_else((df_FR_I$Correct3_5 == "1" & df_FR_I$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_FR_I$Correct3_5 == "0" & df_FR_I$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_FR_I$Correct3_5 == "1" & df_FR_I$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_FR_I$Correct3_5 == "0" & df_FR_I$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_FR_I = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_FR_I



#MC_NoI
PP = sum(if_else((df_MC_NoI$Correct3_5 == "1" & df_MC_NoI$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_MC_NoI$Correct3_5 == "0" & df_MC_NoI$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_MC_NoI$Correct3_5 == "1" & df_MC_NoI$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_MC_NoI$Correct3_5 == "0" & df_MC_NoI$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_MC_NoI = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_MC_NoI



#MC_I
PP = sum(if_else((df_MC_I$Correct3_5 == "1" & df_MC_I$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_MC_I$Correct3_5 == "0" & df_MC_I$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_MC_I$Correct3_5 == "1" & df_MC_I$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_MC_I$Correct3_5 == "0" & df_MC_I$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_MC_I = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_MC_I


```

## McNemar's Test by Question Type
```{r}

# 3.5 versus 4.0

#FR_NoI
mcnemar.test(CT_FR_NoI)
#FR_I
mcnemar.test(CT_FR_I)
#MC_NoI
mcnemar.test(CT_MC_NoI)
#MC_I
mcnemar.test(CT_MC_I)

```



## Contingency Table to Mcnemar's Test by Exam
```{r}

# Filter for Each Question Type
df_ACTM = df %>% filter(Exam == "ACTM") 
df_CAOS = df %>% filter(Exam == "CAOS") 
df_SMUExam = df %>% filter(Exam == "SMUExam") 
df_AP_2011_B = df %>% filter(Exam == "AP2011") 
  


#ACTM
PP = sum(if_else((df_ACTM$Correct3_5 == "1" & df_ACTM$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_ACTM$Correct3_5 == "0" & df_ACTM$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_ACTM$Correct3_5 == "1" & df_ACTM$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_ACTM$Correct3_5 == "0" & df_ACTM$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_ACTM = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_ACTM


#CAOS
PP = sum(if_else((df_CAOS$Correct3_5 == "1" & df_CAOS$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_CAOS$Correct3_5 == "0" & df_CAOS$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_CAOS$Correct3_5 == "1" & df_CAOS$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_CAOS$Correct3_5 == "0" & df_CAOS$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_CAOS = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_CAOS



#SMUExam
PP = sum(if_else((df_SMUExam$Correct3_5 == "1" & df_SMUExam$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_SMUExam$Correct3_5 == "0" & df_SMUExam$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_SMUExam$Correct3_5 == "1" & df_SMUExam$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_SMUExam$Correct3_5 == "0" & df_SMUExam$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_SMUExam = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_SMUExam



#AP_2011_B
PP = sum(if_else((df_AP_2011_B$Correct3_5 == "1" & df_AP_2011_B$Correct4_0 == "1"),1,0))
NP = sum(if_else((df_AP_2011_B$Correct3_5 == "0" & df_AP_2011_B$Correct4_0 == "1"),1,0))
PN = sum(if_else((df_AP_2011_B$Correct3_5 == "1" & df_AP_2011_B$Correct4_0 == "0"),1,0))
NN = sum(if_else((df_AP_2011_B$Correct3_5 == "0" & df_AP_2011_B$Correct4_0 == "0"),1,0))

#McNemar Contingency Table
#Rows 3.5 and Columns 4.0
CT_AP_2011_B = matrix(c(PP,PN,NP,NN),nrow = 2, byrow = TRUE,dimnames = list(c("3.5 Correct", "3.5 Incorrect"), c("4.0 Correct", "4.0 Incorrect")))
CT_AP_2011_B


```

## McNemar's Test by Exam
```{r}

# 3.5 versus 4.0 by Exam

#FR_NoI
mcnemar.test(CT_ACTM)
#FR_I
mcnemar.test(CT_CAOS)
#MC_NoI
mcnemar.test(CT_SMUExam)
#MC_I
mcnemar.test(CT_AP_2011_B)

```

